A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-29 paper/draft.tex
redraft intro

Who benefits from patronage? And what are its consequences for local administration?

Literature on patronage and selection: \citet{robinson2013political}. Extant literature on selection of public officials have primarily focused on who becomes a politician \citet{dal2017becomes}. There remains unanswered questions regarding who enters the bureaucracy, what types of employment they held, and what are the differential compensations among public officials.

@@

One thing is clear: party ties at the electoral level are durable, with most voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}. Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute political factions (\emph{grupos pol\'{i}ticos} ) with well-defined boundaries and power disputes. This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.

@@

Data on partisan affiliation is detailed as well. The \emph{TSE}, the national electoral authority, collects data on all party members in the national territory, including commencement and termination dates, affiliated party, municipality and name of the registered party member, as well as their electoral title.\footnote{The data is available \href{https://filia-consulta.tse.jus.br/}{here}.} Parties are mandated by electoral law to report all party members as well as providing updated registry information throughout the year. Parties have municipal offices in which citizens may update their registry, and these party offices in turn report to the state electoral authority, which relays this information to national electoral authorities. If irregularities are found parties may be sanctioned by the \emph{TSE}.

@@

Finally, we present data on party membership spells in Figure \ref{fig:party_spell}. A few patterns emerge. First, the distribution of spells is rather uniform, with spikes every four years that mirror the jumps in partisanship in Figure \ref{fig:number_partisan}.

\begin{figure}
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_spell.pdf}
    \caption{\textbf{Party memebership spells}. Data includes all party memberships that start in 1985 - the Brazilian democratic transition - and 2019.}
    \label{fig:party_spell}
\end{figure}

@@

In sum, the evidence suggests that party members concentrate in the summits of power. This reinforces priors regarding the control of bureaucratic management by politically appointed bureaucrats in a quantifiable way and in the context of a large, developing country with decentralized structure of local governance \citep{lewis2010politics}. Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it becomes evident that party members are in fact occupying the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.

@@

\section{Empirical strategy}
\label{sec:empirical}

>>>>

Who benefits from patronage? What motivates allocation of public sector jobs in the first place? A

@@

One thing is clear: party ties at the electoral level are durable, with many voters remaining affiliated to a single party for their entire life, as highlighted by figure \emph{FIGURE}. Noting that party affiliations are registered at the municipal level, this empirical fact aligns with qualitative evidence provided by \citet{palmeira1995comicios}, who notes that parties at the local level constitute political factions (\emph{grupos pol\'{i}ticos} ) with well-defined boundaries and power disputes. This relative stability of party ties at the electorate level for the minority of voters who are registered with a parties suggests a distinct dynamic tying an elite group of party members to the city hall.

@@

Data on partisan affiliation in Brazil is rich. The \emph{TSE}, the national electoral authority, collects data on all party members in the national territory, including commencement and termination dates, affiliated party, municipality and name of the registered party member, as well as their electoral title.\footnote{The data is available \href{https://filia-consulta.tse.jus.br/}{here}.} Parties are mandated by electoral law to report all party members as well as providing updated registry information throughout the year. Parties have municipal offices in which citizens may update their registry, and these party offices in turn report to the state electoral authority, which relays this information to national electoral authorities. If irregularities are found parties may be sanctioned by the \emph{TSE}.

@@

Finally, we present data on party membership spells in Figure \ref{fig:party_spell}. A few patterns emerge. First, the distribution of spells is uniform, with spikes every four years that mirror the jumps in partisanship in Figure \ref{fig:number_partisan}. There are no clear patterns with respect to skewedness, suggesting that partisan affiliation duration is distributed in a wave-like pattern, with discontinuities that reflect the occurrence of municipal elections rather than other, long-term dynamics. Overall, the evidence on party membership spells suggests that a constant stream of party members is interspersed with local elections. However, once these membership ties are formed, they are stable and long-lasting.

\begin{figure}
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_spell.pdf}
    \caption{\textbf{Party membership spells}. Data includes all party memberships that start in 1985 - the Brazilian democratic transition - and 2019. Note that there are no significant skewedness in the data either to the left or to the right, indicating that there is a constant stream of party members with the exception of local election spikes.}
    \label{fig:party_spell}
\end{figure}

@@

In sum, the evidence suggests that party members concentrate in the upper echelons of the local bureaucracy. This reinforces priors regarding the control of bureaucratic management by politically appointed bureaucrats in a quantifiable way and in the context of a large, developing country with decentralized structure of local governance \citep{lewis2010politics}. Connecting this empirical evidence with the record that the majority of party members are coming from the upper strata of the formal economy, it is clear that party members are in fact transiting between the upper echelons of the private and public economies, accruing important financial gains in the process and over the long run.

@@

\section{Campaign Finance and the Need for Wealthy Patrons}
Municipal elections are expensive. In 2016, municipal races commanded an average amount of \emph{X} amounts of dollars. In order to finance these races, politicians rely on a variety of resources, and at the local level, these are primarily party-member donations. In particular, while in the larger municipalities there is a greater reliance on corporate donations, for the majority of smaller municipalities, the contribution are in large part individual-level contributions.\

Figure \emph{x} illustrates this pattern.

\section{Empirical strategy}
\label{sec:empirical}
The previous sections present suggestive evidence of systematic differences acros partisan and non-partisan members. In particular, party members seem to be wealthier, less educated and concentrated in the upper strata of the local bureaucracy. However, it is not clear from these aggregate statistics whether the higher compensation is due to

In this section, I perform a simple Oaxaca decomposition to pin down heterogeneous compensation across party and non-partisan members with respect to observables. In particular, I model wages as a function of age, work experience, education level and gender. To allow for greater flexibility with respect to compensation across states, I perform the decomposition broken down by state, and allow coefficients to vary by state, year, and
